---
title: "Analysis of the trends in RD&D spending by energy technology (1974-2022)"
format: html
author: "Disem Sula, Maximilien de La Ruelle"
editor: visual
warning: false
editor_options: 
  chunk_output_type: inline
---

```{r message=FALSE}
#| echo: false
#| message: false
here::i_am("R%D_project.Rproj")
library(here)
library(ggplot2)
library(tidyr)
library(dplyr)
library(vroom)
library(stringr)
library(knitr)
library(patchwork)
```

```{r}
#| message: false
#| echo: false
df<-vroom(here("IEA_Energy_RDD_selected_dataproject.csv"))
gdp<-vroom(here("WorldBank_GDP_data.csv"))

```

### Link to the github repository of this project

[Github repository](https://github.com/max8471/datam_project)

### Sources for the data

For the energy RD&D spending data : The data set is available [here](https://www.iea.org/data-and-statistics/data-product/energy-technology-rd-and-d-budget-database-2#overview) and download it by clicking on "selected data". Delete every sheet of the excel file except "public spending", and delete the first line of this sheet in order to have the years on the top row, then save it as a .csv file.

For the GDP data : Go to this [link](https://databank.worldbank.org/reports.aspx?source=2&series=NY.GDP.MKTP.CD&country#) and select years 1974-2022, then download the data as a .csv file.

### Description of the sources

#### International Energy agency

The International Energy Agency (IEA) is a pivotal organization founded in 1974 within the framework of the Organization for Economic Co-operation and Development (OECD). Comprising major energy consumers and producers globally, the IEA serves to enhance energy security, foster economic growth, and promote environmental sustainability. Its primary functions encompass collecting and analyzing energy data, conducting in-depth studies, and offering policy advice to member countries. The IEA also plays a critical role in coordinating emergency responses during energy supply disruptions.

Key to its mission is advocating for energy efficiency, renewable energy sources, and clean technologies, addressing the challenges of climate change. Through initiatives like the World Energy Outlook reports, the IEA provides comprehensive insights into global energy trends, projections, and policy scenarios. In recent years, the IEA has garnered attention for its commitment to advancing clean energy transitions and contributing to global climate goals. Through collaboration with governments, industries, and international organizations, the IEA actively works towards shaping a sustainable and secure energy future on a global scale.

#### World Bank

The World Bank, established in 1944, stands as a prominent international financial institution dedicated to supporting the development endeavors of nations worldwide. Its fundamental mission revolves around alleviating poverty and promoting sustainable economic development. Comprising two main institutions, the International Bank for Reconstruction and Development (IBRD) and the International Development Association (IDA), the World Bank provides financial assistance in the form of loans, grants, and expert guidance.

The World Bank engages in projects spanning various sectors, including education, health, infrastructure, agriculture, and energy, with a focus on fostering economic growth, reducing inequality, and improving living conditions. In the realm of energy, the World Bank plays a crucial role in financing and implementing projects that enhance access to energy in developing countries, promote renewable energy sources, and improve energy efficiency. Emphasizing sustainable development, environmental protection, and social inclusion, the World Bank's contributions extend beyond financial assistance to actively shaping a more equitable and sustainable global future through comprehensive development initiatives.

### Information about the initial data files

#### IEA data

```{r}
#| message: false
#| echo: false
n_rows_IEA=nrow(df)
n_cols_IEA=ncol(df)
col_names_IEA <- colnames(df)
country_names_IEA <- sapply(unique(df$Country), function(x) iconv(x, "latin1", "UTF-8"))
basic_info_IEA <- data.frame(
  Description = c("Number of Rows", "Number of Columns", "Column Names", "Countries"),
  Value = c(n_rows_IEA, n_cols_IEA, paste(col_names_IEA, collapse=", "), paste(country_names_IEA, collapse = ", ")))
kable(basic_info_IEA)
```

#### GDP data

```{r}
#| message: false
#| echo: false
col_names_GDP <- colnames(gdp)
n_rows_GDP=nrow(gdp)
n_cols_GDP=ncol(gdp)
country_names_GDP=unique(gdp$`Country Name`)
basic_info_GDP <- data.frame(
  Description = c("Number of Rows", "Number of Columns", "Column Names", "Countries"),
  Value = c(n_rows_GDP, n_cols_GDP, paste(col_names_GDP, collapse = ", "), paste(country_names_GDP, collapse = ", ")))
kable(basic_info_GDP)
```

### Filtering the data (IEA)

**Currency Standardization:** All monetary values in the dataset were set to USD using 2022 prices and PPP for uniformity.

```{r}
#| code-fold: true
#Choosing a single currency (USD 2022 prices and PPP), for the whole dataset
filtered_df <- df |>
  filter(Currency == "USD (2022 prices and PPP)")
```

**Timeframe Adjustment:** The dataset was restricted to the years up to 2022, removing data for the year 2023.

```{r}
#| code-fold: true
#Removing year 2023 and stop at year 2022 as for the GDP
filtered_df <- filtered_df[, -ncol(filtered_df)]
```

**Column Removal:** The 'Currency' column in the filtered dataset was removed as it became redundant after currency standardization.

```{r}
#| code-fold: true
#Removing the column Currency in filtered_df since it is not useful anymore
filtered_df <- subset(filtered_df, select = -Currency)
```

### Processing the GDP data to make it correctly mergeable with the IEA one

**Column Naming Standardization:** Only the year information was retained in the column names for GDP

```{r}
#| code-fold: true
#Only keeping the year in the column names of gdp
col_names <- colnames(gdp)

col_names[5:length(col_names)] <- substr(col_names[5:length(col_names)], 1, 4)

colnames(gdp) <- col_names
```

**Country names harmonization:** Three country names were standardized for consistency across datasets.

```{r}
#| code-fold: true
#Modifying 3 names of countries so everything is written with the same standards in both dataframes
gdp <- gdp |> mutate(`Country Name` = ifelse(`Country Name` == "Korea, Rep.", "Korea", `Country Name`))

gdp <- gdp |> mutate(`Country Name` = ifelse(`Country Name` == "Czechia", "Czech Republic", `Country Name`))

filtered_df <- filtered_df |> mutate(`Country` = ifelse(startsWith(`Country`,"Republic of T"), "Turkiye", `Country`))
```

**Filtered GDP Creation:** A new dataframe, 'filtered_gdp,' was created, containing GDP data only for countries present in the IEA dataset.

```{r}
#| code-fold: true
#Keeping the gdp only for the countries that appears in filtered_df, and storing it into a new dataframe named filtered_gdp
distinct_countries <- unique(filtered_df$Country)

filtered_gdp <- gdp[gdp$`Country Name` %in% distinct_countries, ]
```

**Columns Removal:** Columns with irrelevant or redundant information were removed to streamline the dataframe

```{r}
#| code-fold: true
#Now removing the columns that won't be useful for us
filtered_gdp <- subset(filtered_gdp, select = c(-`Series Name`, -`Series Code`, -`Country Code`))
```

**Column Name Adjustment:** The column name 'Country name' was changed to 'Country' for consistency with the naming convention in the IEA filtered dataframe.

```{r}
#| code-fold: true
#Changing the column name to Country from Country name
filtered_gdp <- filtered_gdp |>
  rename(Country = `Country Name`)
```

**Data Type Standardization:** Ensured that all data was in numerical format, facilitating seamless merging of datasets.

```{r, warning=FALSE}
#| code-fold: true
#Making sure all the data is numerical before merging
filtered_df <- filtered_df %>%
  mutate(across(-c(Country, Technology), ~as.numeric(gsub(",", ".", .))))

filtered_gdp <- filtered_gdp %>%
  mutate(across(-`Country`, ~as.numeric(gsub(",", ".", .))))
```

**Unit Standardization:** GDP values were converted to millions to align with the unit of R&D spending, both measured in millions.

```{r}
#| code-fold: true
#GDP in millions (as the R&D spending is in millions)
filtered_gdp <- filtered_gdp |>
  mutate(across(matches("^[12]"), ~./1000000))
```

### Merging the two datasets

**Integration of GDP as a Technology:** A new row was added for each country, treating GDP as a technological variable for comprehensive analysis.

```{r}
#| code-fold: true
#Adding a row for each country to include the GDP into the data as if it is a Technology
combined_df <- bind_rows(filtered_df, filtered_gdp) |>
  arrange(Country, Technology)
options(scipen = 999)
```

**Handling a clarity issue:** Renamed 'NA' entry to 'GDP' in column named "Technology" for clarity and completeness in the dataset

```{r}
#| code-fold: true
#Changing the name of the of the GDP appearing as "NA" to "GDP"
combined_df <- combined_df |>
  mutate(Technology = ifelse(is.na(Technology), "GDP", Technology))
```

## Research question

The global landscape of energy technology and its intersection with economic development have undergone profound transformations over the past five decades. As nations strive for sustainable energy solutions and economic growth, it becomes imperative to investigate the trajectory of energy technology advancements and the corresponding shifts in public Research, Development, and Deployment (RD&D) investments. This research aims to discern patterns, disparities, and potential correlations in the evolution of spending of RD&D on energy technologies among a set of countries over a 50-year span.

**Research Question:** The primary research question guiding this investigation is: "Has there been a discernible change in the energy technology landscape among countries and group of countries such as G7 and EU over the last 50 years, and how does this relate to the proportion of public RD&D spending in relation to their Gross Domestic Product (GDP)?"

The choice of groups like G7 and EU countries for this analysis stems from their global economic significance and shared commitment to addressing energy challenges. By delving into the energy technology trajectories of these nations, we aim to identify trends that may offer insights into the changes in economic structures and innovation ecosystems. Additionally, linking these technological advancements to the proportion of GDP allocated to public RD&D allows to make them comparable from one country to another.

To address the research question, a comprehensive longitudinal analysis will be conducted, spanning five decades from 1974 to 2022. Key variables to be examined include indicators of energy technology evolution, using the public expenditure on RD&D in different technologies to find the adoption of renewable energy sources, advancements in energy efficiency, and breakthroughs in emerging technologies. In parallel, the study will scrutinize each country's GDP and the corresponding public expenditure on RD&D in the energy sector allowing for the assessment of not only technological shifts but also the relative prioritization of energy research within each nation's economic landscape.

The anticipated outcomes of this research include a nuanced understanding of the changing dynamics in energy technology across countries and an exploration of the relationship between economic development and public RD&D spending.

As an introduction, here you will find the evolutions in RD&D spending in 3 major technologies (Nuclear power, fossil fuels and renewables energies) for 3 main areas (Europe, Asia and Oceania, Americas)

```{r}
#| echo: false
filtered_data_IEA_Europe <- combined_df |>
  filter(Country == "IEA Europe" & Technology %in% c("Fossil fuels", "Nuclear", "Renewables"))

filtered_data_IEA_Americas <- combined_df |>
  filter(Country == "IEA Americas" & Technology %in% c("Fossil fuels", "Nuclear", "Renewables"))

filtered_data_IEA_Asia_Oceania <- combined_df |>
  filter(Country == "IEA Asia Oceania" & Technology %in% c("Fossil fuels", "Nuclear", "Renewables"))
```

```{r}
#| echo: false
reshaped_IEA_Europe <- filtered_data_IEA_Europe |>
  pivot_longer(cols = c(starts_with("19"),starts_with("20")), names_to = "Year", values_to = "Value")

reshaped_IEA_Europe$Year <- as.numeric(reshaped_IEA_Europe$Year)

ggplot(reshaped_IEA_Europe, aes(x = Year, y = Value, color = Technology)) +
  geom_line() +
    scale_color_manual(values = c("Renewables" = "darkgreen", "Fossil fuels" = "brown", "Nuclear" = "purple"))+
  labs(title = "Europe - Temporal Trends in Energy RD&D Spending (1974-2022)",
       x = "Year",
       y = "Millions of 2022 USD",
       color = "Technology")




reshaped_IEA_Asia_Oceania <- filtered_data_IEA_Asia_Oceania |>
  pivot_longer(cols = c(starts_with("19"),starts_with("20")), names_to = "Year", values_to = "Value")

reshaped_IEA_Asia_Oceania$Year <- as.numeric(reshaped_IEA_Asia_Oceania$Year)

ggplot(reshaped_IEA_Asia_Oceania, aes(x = Year, y = Value, color = Technology)) +
  geom_line() +
    scale_color_manual(values = c("Renewables" = "darkgreen", "Fossil fuels" = "brown", "Nuclear" = "purple"))+
  labs(title = "Asia and Oceania - Temporal Trends in Energy RD&D Spending (1974-2022)",
       x = "Year",
       y = "Millions of 2022 USD",
       color = "Technology")



reshaped_IEA_Americas <- filtered_data_IEA_Americas |>
  pivot_longer(cols = c(starts_with("19"),starts_with("20")), names_to = "Year", values_to = "Value")

reshaped_IEA_Americas$Year <- as.numeric(reshaped_IEA_Americas$Year)

ggplot(reshaped_IEA_Americas, aes(x = Year, y = Value, color = Technology)) +
  geom_line() +
  scale_color_manual(values = c("Renewables" = "darkgreen", "Fossil fuels" = "brown", "Nuclear" = "purple"))+
  labs(title = "Americas - Temporal Trends in Energy RD&D Spending (1974-2022)",
       x = "Year",
       y = "Millions of 2022 USD",
       color = "Technology")
```

```{r}
#| echo: false
#| message: false
G7 <- c("France", "Italy", "Germany", "United States", "Japan", "United Kingdom", "Canada")
```

```{r}
#| echo: false
#| message: false
reshaped_df <- combined_df |>
  pivot_longer(cols = c(starts_with("19"),starts_with("20")), names_to = "Year", values_to = "Value")

reshaped_df$Year <- as.numeric(reshaped_df$Year)
```

```{r}
#| echo: false
#| #| message: false
g7_summary_df0 <- reshaped_df %>%
  filter(Country %in% G7) %>%
  group_by(Year, Technology) %>%
  summarise(Spending = sum(ifelse(is.na(Value), 0, Value)), .groups = "drop")
```

#Regression G7 countries GDP and Total budget on R&D in energy technologies
```{r}
#| echo: false
#| #| message: false
df_regress <- subset(g7_summary_df0, Technology %in% c('GDP', 'Total Budget'))
df_pivot_regress <- pivot_wider(df_regress, names_from = Technology, values_from = Spending)

#Linear regression
model <- lm(`Total Budget` ~ GDP, data = df_pivot_regress)

summary(model)
```

```{r}
#| echo: false
#| #| message: false
ggplot(df_pivot_regress, aes(x = GDP, y = `Total Budget`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "lightblue") +
  geom_text(aes(label = Year), vjust = -0.5, hjust = -0.5, size = 3) +  # Add Year labels
  labs(title = "Regression Analysis: GDP vs Total Budget",
       x = "GDP",
       y = "Total Budget") +
  theme_minimal()
```

We notice an interesting relationship from the 90's one, so we are going to do a regression limiting the data from the 90's on.

```{r}
#| echo: false
#| message: false
df_regress_90s <- subset(g7_summary_df0, Technology %in% c('GDP', 'Total Budget') & Year >= 1990 & Year <= 2022)
df_pivot_regress_90s <- pivot_wider(df_regress_90s, names_from = Technology, values_from = Spending)

model_subset <- lm(`Total Budget` ~ GDP, data = df_pivot_regress_90s)

summary(model_subset)
```

```{r}
#| echo: false
#| message: false
ggplot(df_pivot_regress_90s, aes(x = GDP, y = `Total Budget`)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "lightblue") +
  geom_text(aes(label = Year), vjust = -0.5, hjust = -0.5, size = 3) +  # Add Year labels
  labs(title = "Regression Analysis: GDP vs Total Budget (1990-2022)",
       x = "GDP",
       y = "Total Budget") +
  theme_minimal()
```

We notice a strong correlation between the GDP and the total budget on energy spending (R-squared=0.76 on the period 1990-2022). To correct this effect, we are going to normalize our values of spending by expressing them as a permillage of the GDP of the country at that time.


## Creation of a data set with values normalized by the GDP


```{r}
#| echo: false
#| message: false
gdp_rows <- reshaped_df %>% filter(Technology == "GDP")

# Values are in %o (permille)
normalized_df <- reshaped_df %>%
  left_join(gdp_rows, by = c("Country", "Year"), suffix = c("", "_GDP")) %>%
  mutate(Value_normalized = Value*1000 / Value_GDP) %>%
  select(-Value_GDP,-Technology_GDP)
```

```{r}
#| echo: false
#| #| message: false
filtered_data_G7 <- normalized_df |>
  filter(Country %in% G7)
filtered_data_G7$Year <- as.numeric(filtered_data_G7$Year)
```

## Now normalizing the same way but bot the G7 grouped data
```{r}
#| echo: false
#| message: false
G7_gdp_rows <- g7_summary_df0 %>% filter(Technology=="GDP")
```


```{r}
#| echo: false
#| message: false
G7_normalized <- g7_summary_df0 %>%
  left_join(G7_gdp_rows,by="Year",suffix=c("","_GDP")) %>%
  mutate(Spending_normalized=Spending*1000/Spending_GDP) %>%
  select(-Technology_GDP,-Spending_GDP)
```



## Comparing spending in %O of the GDP for G7 countries in various technologies

To highlight the interest of the transformation and show how it allows to see the similarities between G7 countries, we have also included the graphs with the spending expressed in 2022 USD instead of per millage of the GDP

```{r}
#| echo: false
#| message: false
#| warning: false
filtered_data_G7_nuclear<- filtered_data_G7 |>
  filter(Country %in% G7 & Technology =="Nuclear")

plot1nuke<-ggplot(filtered_data_G7_nuclear, aes(x = Year, y = Value_normalized, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Nuclear",
       x = "Year",
       y = "%o of GDP",
       color = "Country")

plot2nuke<-ggplot(filtered_data_G7_nuclear, aes(x = Year, y = Value, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Nuclear",
       x = "Year",
       y = "2022 USD",
       color = "Country")
combined_plot_nuke <- plot1nuke + plot2nuke + plot_layout(guides = "collect")&
  theme(legend.position = "bottom")
combined_plot_nuke
```


```{r}
#| echo: false
#| message: false
#| warning: false
filtered_data_G7_renewables<- filtered_data_G7 |>
  filter(Country %in% G7 & Technology =="Renewables")

plot1renew<-ggplot(filtered_data_G7_renewables, aes(x = Year, y = Value_normalized, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Renewables",
       x = "Year",
       y = "%o of GDP",
       color = "Country")

plot2renew<-ggplot(filtered_data_G7_renewables, aes(x = Year, y = Value, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Renewables",
       x = "Year",
       y = "2022 USD",
       color = "Country")
combined_plot_renewables <- plot1renew + plot2renew + plot_layout(guides = "collect")&
  theme(legend.position = "bottom")
combined_plot_renewables
```


```{r}
#| echo: false
#| message: false
#| warning: false
filtered_data_G7_energyefficiency<- filtered_data_G7 |>
  filter(Country %in% G7 & Technology =="Energy efficiency")

plot1nrgeff<-ggplot(filtered_data_G7_energyefficiency, aes(x = Year, y = Value_normalized, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Energy efficiency",
       x = "Year",
       y = "%o of GDP",
       color = "Country")

plot2nrgeff<-ggplot(filtered_data_G7_energyefficiency, aes(x = Year, y = Value, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Energy efficiency",
       x = "Year",
       y = "2022 USD",
       color = "Country")

combined_plot_nrgeff <- plot1nrgeff + plot2nrgeff + plot_layout(guides = "collect")&
  theme(legend.position = "bottom")
combined_plot_nrgeff
```


```{r}
#| echo: false
#| message: false
#| warning: false
filtered_data_G7_fossilfuels<- filtered_data_G7 |>
  filter(Country %in% G7 & Technology =="Fossil fuels")

plot1foss<-ggplot(filtered_data_G7_fossilfuels, aes(x = Year, y = Value_normalized, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Fossil fuels",
       x = "Year",
       y = "%o of GDP",
       color = "Country")

plot2foss<-ggplot(filtered_data_G7_fossilfuels, aes(x = Year, y = Value, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Fossil fuels",
       x = "Year",
       y = "2022 USD",
       color = "Country")

combined_plot_foss <- plot1foss + plot2foss + plot_layout(guides = "collect")&
  theme(legend.position = "bottom")
combined_plot_foss
```


```{r}
#| echo: false
#| message: false
#| warning: false
filtered_data_G7_Hydrogen<- filtered_data_G7 |>
  filter(Country %in% G7 & Technology =="Hydrogen and fuel cells")

plot1hydro<-ggplot(filtered_data_G7_Hydrogen, aes(x = Year, y = Value_normalized, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Hydrogen and fuel cells",
       x = "Year",
       y = "%o of GDP",
       color = "Country")

plot2hydro<-ggplot(filtered_data_G7_Hydrogen, aes(x = Year, y = Value, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Hydrogen and fuel cells",
       x = "Year",
       y = "2022 USD",
       color = "Country")

combined_plot_hydro <- plot1hydro + plot2hydro + plot_layout(guides = "collect")&
  theme(legend.position = "bottom")
combined_plot_hydro
```


```{r}
#| echo: false
#| message: false
#| warning: false
filtered_data_G7_othercrosscutting<- filtered_data_G7 |>
  filter(Country %in% G7 & Technology =="Other cross-cutting technologies/research")

plot1crosscut<-ggplot(filtered_data_G7_othercrosscutting, aes(x = Year, y = Value_normalized, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Other cross-cutting technologies/research",
       x = "Year",
       y = "%o of GDP",
       color = "Country")

plot2crosscut<-ggplot(filtered_data_G7_othercrosscutting, aes(x = Year, y = Value, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Other cross-cutting technologies/research",
       x = "Year",
       y = "2022 USD",
       color = "Country")

combined_plot_crosscut <- plot1crosscut + plot2crosscut + plot_layout(guides = "collect")&
  theme(legend.position = "bottom")
combined_plot_crosscut
```


```{r}
#| echo: false
#| message: false
#| warning: false
filtered_data_G7_Powerandstorage<- filtered_data_G7 |>
  filter(Country %in% G7 & Technology =="Other power and storage technologies")

plot1pow<-ggplot(filtered_data_G7_Powerandstorage, aes(x = Year, y = Value_normalized, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Other power and storage technologies",
       x = "Year",
       y = "%o of GDP",
       color = "Country")

plot2pow<-ggplot(filtered_data_G7_Powerandstorage, aes(x = Year, y = Value, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Other power and storage technologies",
       x = "Year",
       y = "2022 USD",
       color = "Country")

combined_plot_pow <- plot1pow + plot2pow + plot_layout(guides = "collect")&
  theme(legend.position = "bottom")
combined_plot_pow
```


```{r}
#| echo: false
#| message: false
#| warning: false
filtered_data_G7_totalbudget<- filtered_data_G7 |>
  filter(Country %in% G7 & Technology =="Total Budget")

plot1tot<-ggplot(filtered_data_G7_totalbudget, aes(x = Year, y = Value_normalized, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Total Budget",
       x = "Year",
       y = "%o of GDP",
       color = "Country")

plot2tot<-ggplot(filtered_data_G7_totalbudget, aes(x = Year, y = Value, group=Country, color=Country)) +
  geom_line() +
  labs(title = "G7 - Total Budget",
       x = "Year",
       y = "2022 USD",
       color = "Country")

combined_plot_tot <- plot1tot + plot2tot + plot_layout(guides = "collect")&
  theme(legend.position = "bottom")
combined_plot_tot
```


We see that for quite every technology, the spending in public rd&d, when expressed as a per millage of the GDP, is very similar and follows the same trend from one G7 country to another. And even more on the period 1990-2022, where the correlation between GDP and Total budget has been highlighted precedently.
This will allow us to analyze the trend in spending at the G7 level while staying relevant, since this group is homogeneous enough.

```{r}
#| echo: false
#| message: false
filtered_G7 <- G7_normalized |>
  filter(Technology %in% c("Fossil fuels", "Nuclear", "Renewables","Energy efficiency","Hydrogen and fuel cells","Other cross-cutting technologies/research","Other power and storage technologies"))

plot1g7tech<-ggplot(filtered_G7, aes(x = Year, y = Spending_normalized, color = Technology)) +
  geom_line()+
  labs(title = "Spending on various energy technologies for G7 (1974-2022)",
       x = "Year",
       y = "Spending (per millage of the GDP)",
       color = "Technology")

G7_wonuc <- G7_normalized |>
  filter(Technology %in% c("Fossil fuels", "Renewables","Energy efficiency","Hydrogen and fuel cells","Other cross-cutting technologies/research","Other power and storage technologies"))


plot2g7tech<-ggplot(G7_wonuc, aes(x = Year, y = Spending_normalized, color = Technology)) +
  geom_line(show.legend=FALSE) +
  labs(title = "Nuclear excluded",
       x = "Year",
       y = "Spending (per millage of the GDP)",
       color = "Technology")

combined_plot_g7tech <- (plot1g7tech + plot2g7tech) + plot_layout(guides = "collect")&
  theme(legend.position = "bottom")
combined_plot_g7tech
```
